---
title: "Rushing"
author: "Richard Charles"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(stats)
library(factoextra)
library(sail)
library(caret)
library(leaps)
library(glmnet)
library(tree)
library(gbm)
library(randomForest)
```

```{r}
set.seed(445)
master <- read.csv('./Data/masterDF[2018-2020].csv')
rush_data <- master[, c(14:17, 24:24, 45:46)]
head(rush_data)
```

```{r}
train_data <- rush_data[which(rush_data$RushYds > 0 & !is.na(rush_data$RushLng)), -c(8:8)]
N <- nrow(train_data)
K <- 10
folds <- sample(1:K, N, replace = TRUE)
for (i in 1:N) {
  if (is.na(train_data$FirstDRush[i])) {
    train_data$FirstDRush[i] <- 0
  }
  if (is.na(train_data$RushAtt[i])) {
    train_data$RushAtt[i] <- 0
  }
}
head(train_data)
```

```{r}
ggpairs(train_data[,c(1:6, 7)])
```

```{r}
summary(lm(RushAtt ~ . - RushYds - RushTD - RushLng - FL - FirstDRush - RushYBC, data = train_data))
```

```{r}
summary(lm(RushYds ~ . - RushAtt - RushTD - RushLng - FL - FirstDRush - RushYBC, data = train_data))
```

```{r}
summary(lm(RushTD ~ . - RushYds - RushAtt - RushLng - FL - FirstDRush - RushYBC, data = train_data))
```

```{r}
summary(lm(RushLng ~ . - RushYds - RushTD - RushAtt - FL - FirstDRush - RushYBC, data = train_data))
```

```{r}
summary(lm(FL ~ . - RushYds - RushTD - RushLng - RushAtt - FirstDRush - RushYBC, data = train_data))
```

```{r}
summary(lm(FirstDRush ~ . - RushYds - RushTD - RushLng - FL - RushAtt - RushYBC, data = train_data))
```

```{r}
summary(lm(RushYBC ~ . - RushYds - RushTD - RushLng - FL - RushAtt - FirstDRush, data = train_data))
```

```{r}
res = data_frame()
n = nrow(train_data)
folds = sample(seq_len(10),n,replace = TRUE)
k <- 10
for(i in seq_len(k)) {
  trn = folds != i
  m0 <- lm(RushAtt ~ . - RushYds - RushTD - RushLng - FL, data = train_data)
  m1 <- lm(RushYds ~ . - RushAtt - RushTD - RushLng - FL, data = train_data)
  m2 <- lm(RushTD ~ . - RushYds - RushAtt - RushLng - FL, data = train_data)
  m3 <- lm(FL ~ . - RushYds - RushTD - RushLng - RushAtt, data = train_data)

  p0 <- predict(m0, train_data[-trn,])
  p1 <- predict(m1, train_data[-trn,])
  p2 <- predict(m2, train_data[-trn,])
  p3 <- predict(m3, train_data[-trn,])
  
  true_rushatt = train_data[-trn, ]$RushAtt
  true_rushyds = train_data[-trn, ]$RushYds
  true_rushtd = train_data[-trn, ]$RushTD
  true_fl = train_data[-trn, ]$FL
  
  data.frame(terms = 2, model = "aRush Att", true = true_rushatt, pred = p0) %>%
    bind_rows(data_frame(terms = 3, model = "cRush TD's", true = true_rushtd, pred = p2)) %>%
    bind_rows(data_frame(terms = 4, model = "bRush Yards", true = true_rushyds, pred = p1)) %>%
    bind_rows(data_frame(terms = 5, model = "dFumbles Lost", true = true_fl, pred = p3)) %>%
    mutate(mse = (true - pred)^2) %>%
    group_by(model) %>%
    summarise(mse = mean(mse)) -> test_mse_k
  
  res %>% bind_rows(test_mse_k) -> res
}

summary(m0)
mlr_results = res %>%
  group_by(model) %>%
  summarise(MSE = mean(mse))
mlr_results = mlr_results %>%
  arrange("Rush Att", "Rush Yards", "Rush TD's", "Fumbles Lost" )
mlr_results
```

# LASSO Regression

```{r}
x <- model.matrix(RushAtt ~ . - RushYds - RushTD - RushLng - FL, data = train_data)
y <- train_data$RushAtt
lasso_model <- cv.glmnet(x, y, type.measure = 'mse', nfolds = 10); lasso_model
plot(lasso_model)
lasso_model <- glmnet(x, y, lambda = exp(-2)); lasso_model
coef(lasso_model)
lasso_mse <- rep(NA, K)
for (k in 1:K) {
  trn <- train_data[folds != k,]
  vld <- train_data[folds == k,]
  
  x <- model.matrix(RushAtt ~ . - RushYds - RushTD - RushLng - FL, trn)
  y <- trn$RushAtt
  
  lasso_model <- glmnet(x, y, lambda = exp(-2))
  
  pred_x <- model.matrix(RushAtt ~ . - RushYds - RushTD - RushLng - FL, vld)
  pred <- predict(lasso_model, pred_x)
  
  lasso_mse[k] <- mean((vld$RushAtt - pred)^2)
}
recMSE <- round(mean(lasso_mse), 3)
```

```{r}
x <- model.matrix(RushYds ~ . - RushAtt - RushTD - RushLng - FL, data = train_data)
y <- train_data$RushYds
lasso_model <- cv.glmnet(x, y, type.measure = 'mse', nfolds = 10); lasso_model
plot(lasso_model)
lasso_model <- glmnet(x, y, lambda = exp(1)); lasso_model
coef(lasso_model)
lasso_mse <- rep(NA, K)
for (k in 1:K) {
  trn <- train_data[folds != k,]
  vld <- train_data[folds == k,]
  
  x <- model.matrix(RushYds ~ . - RushAtt - RushTD - RushLng - FL, trn)
  y <- trn$RushYds
  
  lasso_model <- glmnet(x, y, lambda = exp(1))
  
  pred_x <- model.matrix(RushYds ~ . - RushAtt - RushTD - RushLng - FL, vld)
  pred <- predict(lasso_model, pred_x)
  
  lasso_mse[k] <- mean((vld$RushYds - pred)^2)
}
recYdsMSE <- round(mean(lasso_mse), 3)
```


```{r}
x <- model.matrix(RushTD ~ . - RushYds - RushAtt - RushLng - FL, data = train_data)
y <- train_data$RushTD
lasso_model <- cv.glmnet(x, y, type.measure = 'mse', nfolds = 10); lasso_model
plot(lasso_model)
lambda <- exp
lasso_model <- glmnet(x, y, lambda = exp(-5)); lasso_model
coef(lasso_model)
lasso_mse <- rep(NA, K)
for (k in 1:K) {
  trn <- train_data[folds != k,]
  vld <- train_data[folds == k,]
  
  x <- model.matrix(RushTD ~ . - RushYds - RushAtt - RushLng - FL, trn)
  y <- trn$RushTD
  
  lasso_model <- glmnet(x, y, lambda = exp(-5))
  
  pred_x <- model.matrix(RushTD ~ . - RushYds - RushAtt - RushLng - FL, vld)
  pred <- predict(lasso_model, pred_x)
  
  lasso_mse[k] <- mean((vld$RushTD - pred)^2)
}
recTDMSE <- round(mean(lasso_mse), 3)
```


```{r}
x <- model.matrix(FL ~ . - RushYds - RushTD - RushLng - RushAtt, train_data)
y <- train_data$FL
lasso_model <- cv.glmnet(x, y, type.measure = 'mse', nfolds = 10); lasso_model
plot(lasso_model)
lasso_model <- glmnet(x, y, lambda = exp(-3)); lasso_model
coef(lasso_model)
lasso_mse <- rep(NA, K)
for (k in 1:K) {
  trn <- train_data[folds != k,]
  vld <- train_data[folds == k,]
  
  x <- model.matrix(FL ~ . - RushYds - RushTD - RushLng - RushAtt, trn)
  y <- trn$FL
  
  lasso_model <- glmnet(x, y, lambda = exp(-3))
  
  pred_x <- model.matrix(FL ~ . - RushYds - RushTD - RushLng - RushAtt, vld)
  pred <- predict(lasso_model, pred_x)
  
  lasso_mse[k] <- mean((vld$FL - pred)^2)
}
FLMSE <- round(mean(lasso_mse), 3)
```

10-fold cross validation LASSO regression MSEs:
```{r}
lasso_results <- data.frame(Model = c('RushAtt', 'RushYds', 'RushTD', 'FL'), MSE = rep(NA, 4))
lasso_results$MSE[which(lasso_results$Model == 'RushAtt')] <- recMSE
lasso_results$MSE[which(lasso_results$Model == 'RushYds')] <- recYdsMSE
lasso_results$MSE[which(lasso_results$Model == 'RushTD')] <- recTDMSE
lasso_results$MSE[which(lasso_results$Model == 'FL')] <- FLMSE
lasso_results
```

PCR

```{r}
pca_model <- prcomp(train_data[,c(1:3, 5)], center = TRUE, scale = TRUE)
summary(pca_model)
```
Percentage of total variance explained by each principal component. 
```{r}
fviz_eig(pca_model)
```
Principal component loading matrix. 
```{r}
pca_scores <- data.frame(pca_model$x)
data.frame(pca_model$rotation[, 1:4])
```
Principal component biplot for PC1 and PC2.
```{r}
biplot(pca_model)
```

Principal components regression performed on the `RushAtt` variable. 
```{r}
pca_train <- pca_scores %>%
  mutate(RushAtt = train_data$RushAtt)
results <- data.frame(pc = 1:5, mse = rep(NA, 5))
for (pc in results$pc) {
  mse <- rep(NA, K)
  for (k in 1:K) {
    trn <- pca_train[folds != k, c(1:pc, 5)]
    vld <- pca_train[folds == k, c(1:pc, 5)]
    
    pca_model <- lm(RushAtt ~ ., data = trn)
    pred <- predict(pca_model, vld)
    mse[k] <- mean((vld$RushAtt - pred)^2)
  }
  results$mse[which(results$pc == pc)] <- mean(mse)
}
ggplot(results) +
  geom_line(aes(pc, mse)) +
  geom_point(aes(4, mse[4]), color = 'red') +
  labs(title='10-Fold Cross Validation MSE', x='Principal Components', y='10-Fold MSE')
recMSE <- results$mse[which(results$pc == 4)]
```
Principal components regression performed on the `RushYds` variable. 
```{r}
pca_train <- pca_scores %>%
  mutate(RushYds = train_data$RushYds)
results <- data.frame(pc = 1:5, mse = rep(NA, 5))
for (pc in results$pc) {
  mse <- rep(NA, K)
  for (k in 1:K) {
    trn <- pca_train[folds != k, c(1:pc, 5)]
    vld <- pca_train[folds == k, c(1:pc, 5)]
    
    pca_model <- lm(RushYds ~ ., data = trn)
    pred <- predict(pca_model, vld)
    mse[k] <- mean((vld$RushYds - pred)^2)
  }
  results$mse[which(results$pc == pc)] <- mean(mse)
}
ggplot(results) +
  geom_line(aes(pc, mse)) +
  geom_point(aes(4, mse[4]), color = 'red') +
  labs(title='10-Fold Cross Validation MSE', x='Principal Components', y='10-Fold MSE')
recYdsMSE <- results$mse[which(results$pc == 4)]
```
Principal components regression performed on the `RushTD` variable. 
```{r}
pca_train <- pca_scores %>%
  mutate(RushTD = train_data$RushTD)
results <- data.frame(pc = 1:5, mse = rep(NA, 5))
for (pc in results$pc) {
  mse <- rep(NA, K)
  for (k in 1:K) {
    trn <- pca_train[folds != k, c(1:pc, 5)]
    vld <- pca_train[folds == k, c(1:pc, 5)]
    
    pca_model <- lm(RushTD ~ ., data = trn)
    pred <- predict(pca_model, vld)
    mse[k] <- mean((vld$RushTD - pred)^2)
  }
  results$mse[which(results$pc == pc)] <- mean(mse)
}
ggplot(results) +
  geom_line(aes(pc, mse)) +
  geom_point(aes(3, mse[3]), color = 'red') +
  labs(title='10-Fold Cross Validation MSE', x='Principal Components', y='10-Fold MSE')
recTDMSE <- results$mse[which(results$pc == 3)]
```
Principal components regression performed on the `FL` variable. 
```{r}
pca_train <- pca_scores %>%
  mutate(FL = train_data$FL)
results <- data.frame(pc = 1:5, mse = rep(NA, 5))
for (pc in results$pc) {
  mse <- rep(NA, K)
  for (k in 1:K) {
    trn <- pca_train[folds != k, c(1:pc, 5)]
    vld <- pca_train[folds == k, c(1:pc, 5)]
    
    pca_model <- lm(FL ~ ., data = trn)
    pred <- predict(pca_model, vld)
    mse[k] <- mean((vld$FL - pred)^2)
  }
  results$mse[which(results$pc == pc)] <- mean(mse)
}
ggplot(results) +
  geom_line(aes(pc, mse)) +
  geom_point(aes(2, mse[2]), color = 'red') +
  labs(title='10-Fold Cross Validation MSE', x='Principal Components', y='10-Fold MSE')
FLMSE <- results$mse[which(results$pc == 2)]
```

10-fold cross validation PC regression MSEs:
```{r}
pcr_results <- data.frame(Model = c('RushAtt', 'RushYds', 'RushTD', 'FL'), MSE = rep(NA, 4))
pcr_results$MSE[which(pcr_results$Model == 'RushAtt')] <- recMSE
pcr_results$MSE[which(pcr_results$Model == 'RushYds')] <- recYdsMSE
pcr_results$MSE[which(pcr_results$Model == 'RushTD')] <- recTDMSE
pcr_results$MSE[which(pcr_results$Model == 'FL')] <- FLMSE
pcr_results
```

```{r, warning=FALSE, message=FALSE}
results <- data.frame(lambda = seq(0, .5, length.out = 5), MSE = rep(NA, 5))
for (lambda in results$lambda) {
  mse <- rep(NA, K)
  
  for (k in 1:K) {
    trn <- train_data[folds != k,]
    vld <- train_data[folds == k,]
    
    model <- gbm(RushAtt ~ . - RushYds - RushTD - RushLng - FL, distribution = 'gaussian', n.trees = 100, shrinkage = lambda, trn)
    
    pred <- round(predict(model, vld))
    
    mse[k] <- mean((vld$RushAtt - pred)^2)
  }
  
  results$MSE[which(results$lambda == lambda)] <- mean(mse)
}
ggplot(results) +
  geom_line(aes(lambda, MSE)) +
  geom_point(aes(.125, results$MSE[which(results$lambda == .125)]), color = 'red') +
  labs(title='Shrinkage Factor Selection', subtitle='10-fold CV MSE')
recMSE <- round(results$MSE[which(results$MSE == min(results$MSE))], 3)
```
Boosted tree model to predict `RushYds` variable.
```{r, warning=FALSE, message=FALSE}
results <- data.frame(lambda = seq(0, .5, length.out = 5), MSE = rep(NA, 5))
for (lambda in results$lambda) {
  mse <- rep(NA, K)
  
  for (k in 1:K) {
    trn <- train_data[folds != k,]
    vld <- train_data[folds == k,]
    
    model <- gbm(RushYds ~ . - RushAtt - RushTD - RushLng - FL, data = trn, distribution = 'gaussian', n.trees = 100, shrinkage = lambda)
    
    pred <- round(predict(model, vld))
    
    mse[k] <- mean((vld$RushYds - pred)^2)
  }
  
  results$MSE[which(results$lambda == lambda)] <- mean(mse)
}
ggplot(results) +
  geom_line(aes(lambda, MSE)) +
  geom_point(aes(.125, results$MSE[which(results$lambda == .125)]), color = 'red') +
  labs(title='Shrinkage Factor Selection', subtitle='10-fold CV MSE')
recYdsMSE <- round(results$MSE[which(results$MSE == min(results$MSE))], 3)
```
Boosted tree model to predict `RushTD` variable.
```{r, warning=FALSE, message=FALSE}
results <- data.frame(lambda = seq(0, .5, length.out = 5), MSE = rep(NA, 5))
for (lambda in results$lambda) {
  mse <- rep(NA, K)
  
  for (k in 1:K) {
    trn <- train_data[folds != k,]
    vld <- train_data[folds == k,]
    
    model <- gbm(RushTD ~ . - RushYds - RushAtt - RushLng - FL, data = trn, distribution = 'gaussian', n.trees = 100, shrinkage = lambda)
    
    pred <- round(predict(model, vld))
    
    mse[k] <- mean((vld$RushTD - pred)^2)
  }
  
  results$MSE[which(results$lambda == lambda)] <- mean(mse)
}
ggplot(results) +
  geom_line(aes(lambda, MSE)) +
  geom_point(aes(.125, results$MSE[which(results$lambda == .125)]), color = 'red') +
  labs(title='Shrinkage Factor Selection', subtitle='10-fold CV MSE')
recTDMSE <- round(results$MSE[which(results$MSE == min(results$MSE))], 3)
```
Boosted tree model to predict `FL` variable.
```{r, warning=FALSE, message=FALSE}
results <- data.frame(lambda = seq(0, .5, length.out = 5), MSE = rep(NA, 5))
for (lambda in results$lambda) {
  mse <- rep(NA, K)
  
  for (k in 1:K) {
    trn <- train_data[folds != k,]
    vld <- train_data[folds == k,]
    
    model <- gbm(FL ~ . - RushAtt - RushYds - RushTD, data = trn, distribution = 'gaussian', n.trees = 100, shrinkage = lambda)
    
    pred <- round(predict(model, vld))
    
    mse[k] <- mean((vld$FL - pred)^2)
  }
  
  results$MSE[which(results$lambda == lambda)] <- mean(mse)
}
ggplot(results) +
  geom_line(aes(lambda, MSE)) +
  geom_point(aes(.125, results$MSE[which(results$lambda == .125)]), color = 'red') +
  labs(title='Shrinkage Factor Selection', subtitle='10-fold CV MSE')
FLMSE <- round(results$MSE[which(results$MSE == min(results$MSE))], 3)
```
10-fold cross validation boosted tree MSEs:
```{r}
boost_results <- data.frame(Model = c('RushAtt', 'RushYds', 'RushTD', 'FL'), MSE = rep(NA, 4))
boost_results$MSE[which(boost_results$Model == 'RushAtt')] <- recMSE
boost_results$MSE[which(boost_results$Model == 'RushYds')] <- recYdsMSE
boost_results$MSE[which(boost_results$Model == 'RushTD')] <- recTDMSE
boost_results$MSE[which(boost_results$Model == 'FL')] <- FLMSE
boost_results
```

# Bagging

```{r}
mse <- rep(NA, K)
for (k in 1:K) {
  trn <- train_data[folds != k,]
  vld <- train_data[folds == k,]
  model <- randomForest(RushAtt ~ . - RushYds - RushTD - FL, data = trn, mtry = 15, ntree = 10)
  pred <- round(predict(model, vld))
  mse[k] <- mean((vld$RushAtt - pred)^2)
}
recMSE <- round(mean(mse), 3)
```

```{r}
mse <- rep(NA, K)
for (k in 1:K) {
  trn <- train_data[folds != k,]
  vld <- train_data[folds == k,]
  model <- randomForest(RushYds ~ . - RushAtt - RushTD - FL, data = trn, mtry = 15, ntree = 10)
  pred <- round(predict(model, vld))
  mse[k] <- mean((vld$RushYds - pred)^2)
}
recYdsMSE <- round(mean(mse), 3)
```

```{r, warning=FALSE}
mse <- rep(NA, K)
for (k in 1:K) {
  trn <- train_data[folds != k,]
  vld <- train_data[folds == k,]
  model <- randomForest(RushTD ~ . - RushAtt - RushYds - FL, data = trn, mtry = 15, ntree = 10)
  pred <- round(predict(model, vld))
  mse[k] <- mean((vld$RushTD - pred)^2)
}
recTDMSE <- round(mean(mse), 3)
```

```{r, warning=FALSE}
mse <- rep(NA, K)
for (k in 1:K) {
  trn <- train_data[folds != k,]
  vld <- train_data[folds == k,]
  model <- randomForest(FL ~ . - RushAtt - RushYds - RushTD - FL, data = trn, mtry = 15, ntree = 10)
  pred <- round(predict(model, vld))
  mse[k] <- mean((vld$FL - pred)^2)
}
FLMSE <- round(mean(mse), 3)
```

10-fold cross validation bagged tree MSEs:
```{r}
bag_results <- data.frame(Model = c('RushAtt', 'RushYds', 'RushTD', 'FL'), MSE = rep(NA, 4))
bag_results$MSE[which(bag_results$Model == 'RushAtt')] <- recMSE
bag_results$MSE[which(bag_results$Model == 'RushYds')] <- recYdsMSE
bag_results$MSE[which(bag_results$Model == 'RushTD')] <- recTDMSE
bag_results$MSE[which(bag_results$Model == 'FL')] <- FLMSE
bag_results
```

```{r}
results <- data.frame(Model = c('RushAtt', 'RushYds', 'RushTD', 'FL'),
                      MLR = mlr_results$MSE,
                      LASSO = lasso_results$MSE,
                      PCR = pcr_results$MSE,
                      Boost = boost_results$MSE,
                      Bag = bag_results$MSE); results

```

