---
title: "Scraping Script"
author: "Adam Kiehl"
date: "11/10/21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(XML)
```

## Variable Setup

A dictionary relating full team names to their three-letter codes. 
```{r}
teams <- data.frame(team = character(), team_code = character()) %>%
  rbind(data.frame(team = 'Arizona Cardinals', team_code = 'crd')) %>%
  rbind(data.frame(team = 'Atlanta Falcons', team_code = 'atl')) %>%
  rbind(data.frame(team = 'Baltimore Ravens', team_code = 'rav')) %>%
  rbind(data.frame(team = 'Buffalo Bills', team_code = 'buf')) %>%
  rbind(data.frame(team = 'Carolina Panthers', team_code = 'car')) %>%
  rbind(data.frame(team = 'Chicago Bears', team_code = 'chi')) %>%
  rbind(data.frame(team = 'Cincinnati Bengals', team_code = 'cin')) %>%
  rbind(data.frame(team = 'Cleveland Browns', team_code = 'cle')) %>%
  rbind(data.frame(team = 'Dallas Cowboys', team_code = 'dal')) %>%
  rbind(data.frame(team = 'Denver Broncos', team_code = 'den')) %>%
  rbind(data.frame(team = 'Detroit Lions', team_code = 'det')) %>%
  rbind(data.frame(team = 'Green Bay Packers', team_code = 'gnb')) %>%
  rbind(data.frame(team = 'Houston Texans', team_code = 'htx')) %>%
  rbind(data.frame(team = 'Indianapolis Colts', team_code = 'clt')) %>%
  rbind(data.frame(team = 'Jacksonville Jaguars', team_code = 'jax')) %>%
  rbind(data.frame(team = 'Kansas City Chiefs', team_code = 'kan')) %>%
  rbind(data.frame(team = 'Las Vegas Raiders', team_code = 'rai')) %>%
  rbind(data.frame(team = 'Los Angeles Chargers', team_code = 'sdg')) %>%
  rbind(data.frame(team = 'Los Angeles Rams', team_code = 'ram')) %>%
  rbind(data.frame(team = 'Miami Dolphins', team_code = 'mia')) %>%
  rbind(data.frame(team = 'Minnesota Vikings', team_code = 'min')) %>%
  rbind(data.frame(team = 'New England Patriots', team_code = 'nwe')) %>%
  rbind(data.frame(team = 'New Orleans Saints', team_code = 'nor')) %>%
  rbind(data.frame(team = 'New York Giants', team_code = 'nyg')) %>%
  rbind(data.frame(team = 'New York Jets', team_code = 'nyj')) %>%
  rbind(data.frame(team = 'Oakland Raiders', team_code = 'rai')) %>%
  rbind(data.frame(team = 'Philadelphia Eagles', team_code = 'phi')) %>%
  rbind(data.frame(team = 'Pittsburgh Steelers', team_code = 'pit')) %>%
  rbind(data.frame(team = 'San Diego Chargers', team_code = 'sdg')) %>%
  rbind(data.frame(team = 'San Francisco 49ers', team_code = 'sfo')) %>%
  rbind(data.frame(team = 'Seattle Seahawks', team_code = 'sea')) %>%
  rbind(data.frame(team = 'St. Louis Rams', team_code = 'ram')) %>%
  rbind(data.frame(team = 'Tampa Bay Buccaneers', team_code = 'tam')) %>%
  rbind(data.frame(team = 'Tennessee Titans', team_code = 'oti')) %>%
  rbind(data.frame(team = 'Washington Football Team', team_code = 'was'))
```

A vector of years of interest to scrape. This is passed into the scraping functions.
```{r}
years <- as.character(2020:2020)
```

Standardized variable names for each table of interest and merged master table. These are passed into the scraping functions. 
```{r}
names <- list()
names[['player_offense']] <- c('Player', 'Team', 'Cmp', 'PassAtt', 'PassYds', 'PassTD', 'PassInt', 'Sk', 'YdsLost', 'PassLng', 'Rate',
                               'RushAtt', 'RushYds', 'RushTD', 'RushLng', 'Tgt', 'Rec', 'RecYds', 'RecTD', 'RecLng', 'Fmb', 'FL')
names[['passing_advanced']] <- c('Player', 'Team', 'Cmp', 'PassAtt', 'PassYds', 'FirstDPass', 'FirstDPassPer', 'IAY', 'IAYPerPA', 'CAY', 
                                 'CAYPerCmp', 'CAYPerPA', 'PassYAC', 'YACPerCmp', 'PassDrops', 'DropPer', 'BadThrow', 'BadPer', 'Sk', 'Bltz', 
                                 'Hrry', 'Hits', 'Prss', 'PrssPer', 'Scrm', 'YdsPerScr')
names[['rushing_advanced']] <- c('Player', 'Team', 'RushAtt', 'RushYds', 'FirstDRush', 'RushYBC', 'YBCPerAtt', 'RushYAC', 'YACPerAtt', 
                                 'RushBrkTkl', 'AttPerBr')
names[['receiving_advanced']] <- c('Player', 'Team', 'Tgt', 'Rec', 'RecYds', 'RecTD', 'FirstDRec', 'RecYBC', 'YBCPerR', 'RecYAC', 'YACPerR',
                                   'ADOT', 'RecBrkTkl', 'RecPerBr', 'RecDrop', 'DropPerRec', 'RecInt', 'Rat')
names[['master']] <- c('Player', 'Game', 'Team', 'Cmp', 'PassAtt', 'PassYds', 'PassTD', 'PassInt', 'Sk', 'YdsLost', 'PassLng', 'Rate',
                       'RushAtt', 'RushYds', 'RushTD', 'RushLng', 'Tgt', 'Rec', 'RecYds', 'RecTD', 'RecLng', 'Fmb', 'FL', 'FirstDPass',
                       'FirstDPassPer', 'IAY', 'IAYPerPA', 'CAY', 'CAYPerCmp', 'CAYPerPA', 'PassYAC', 'YACPerCmp', 'PassDrops', 'DropPer',
                       'BadThrow', 'BadPer', 'Bltz', 'Hrry', 'Hits', 'Prss', 'PrssPer', 'Scrm', 'YdsPerScr', 'FirstDRush', 'RushYBC', 
                       'YBCPerAtt', 'RushYAC', 'YACPerAtt', 'RushBrkTkl', 'AttPerBr', 'FirstDRec', 'RecYBC', 'YBCPerR', 'RecYAC', 'YACPerR', 
                       'ADOT', 'RecBrkTkl', 'RecPerBr', 'RecDrop', 'DropPerRec', 'RecInt', 'Rat')
```

## Function Setup

scrape_stats() is the parent function used to scrape all stats of interest. Given a vector of years, it loops through all (year, team) combinations to find a set of all unique games to scrape. The tables of interest are `player_offense`, `passing_advanced`, `rushing_advanced`, and `receiving_advanced`. This function will return a large (62 columns) data frame of all unique (player, game) observations for the given period. Helper functions include scrape_games() to find unique games and scrape_tables() to extract data from selected game pages. 
```{r}
scrape_stats <- function(years, names) {
  # Begin a timer to track elapsed progress
  ptm <- proc.time()
  
  # Use master game codes list to track already-explored games
  master_games <- c()
  # Build temporary data frame to fill
  temp <- data.frame(matrix(ncol = length(names[['master']]), nrow = 0))
  names(temp) <- names[['master']]
  
  # Loop through years
  for (year in years) { 
    # Loop through teams
    for (team in unique(teams$team_code)) { 
      # Print progress
      elapse <- proc.time() - ptm
      elapse <- round(as.numeric(elapse[3]), 1)
      print(paste('Scraping (', year, ', ', team, ');  Time Elapsed: ', elapse, ' sec', sep = ''))
      
      # Retrieve unexplored game codes for (year, team) combinations using helper functions
      game_codes <- scrape_games(team, year, master_games)
      
      # Loop through game codes
      for (game_code in game_codes) {
        # Scrape tables and return merged version with helper function 
        temp <- rbind(temp, scrape_tables(game_code, names))
      }
    }
    print('')
  }
  
  # Report total scraping time
  elapse <- proc.time() - ptm
  elapse <- round(as.numeric(elapse[3]), 1)
  print(paste('Total Time Elapsed: ', elapse, ' sec', sep = ''))
  
  return(temp)
}
```

A helper function for scrape_stats(). Given a team code and a year, this function returns the game codes of all regular season games played. 
```{r, warning=FALSE}
scrape_games <- function(team, year, master_games) {
  # Extract game log for (team, year) combination
  url <- paste('https://www.pro-football-reference.com/teams/', team, '/', year, '/gamelog/', sep = '')
  page <- read_html(url)
  
  # Extract tables and parse game log
  tbls <- page %>%
    html_nodes('table')
  for (tbl in tbls) {
    id <- html_attr(tbl, 'id')
    if (id == paste('gamelog', year, sep = '')) {
      df <- html_table(tbl) %>%
        as.data.frame()
      
      # Rename fields
      names(df) <- df[1,]
      names(df)[7] <- 'HA'
      df <- df[-1,]
      
      # Return constructed game codes
      codes <- build_game_codes(df$Date, df$HA, df$Opp, team, year)
      codes <- codes[!(codes %in% master_games)]
      return(codes)
    }
  }
}
```

A helper function for scrape_games(). This function takes scraped date and opponent data from a team's game log and constructs game codes according to PFR's naming system. The format is yyyymmdd0{home team code}. 
```{r}
build_game_codes <- function(dates, HAs, opps, team, year) {
  # Format as date and remove default year
  codes <- dates %>%
    as.Date(format = '%B%d') %>%
    as.character()  %>%
    substr(6, 10)  %>%
    str_replace('-', '')

  # Loop through games
  for (i in 1:length(codes)) {
    # Determine whether year or year+1 should be used for the game code and append
    if (as.numeric(substr(codes[i], 1, 2)) < 6) {
      codes[i] <- paste(as.character(as.numeric(year) + 1), codes[i], '0', sep = '')
    }
    else{
      codes[i] <- paste(year, codes[i], '0', sep = '')
    }
    
    # Determine home team and append
    if (HAs[i] == '@') {
      codes[i] <- paste(codes[i], decode_team(opps[i]), sep = '')
    }
    else {
      codes[i] <- paste(codes[i], team, sep = '')
    }
  }
  
  # Return constructed game codes
  return(codes)
}
```

A simple helper function for build_game_codes() that translates a full team name into its respective team code. 
```{r}
decode_team <- function(team) {
  teams$team_code[which(teams$team == team)]
}
```

A helper function for scrape_stats(). Given a game code, this function accesses the associated game page on PFR and extracts the relevant data. A merged version of the data tables is returned. 
```{r}
scrape_tables <- function(game_code, names) {
  # Construct URL and extract tables into list
  url <- paste('https://www.pro-football-reference.com/boxscores/', game_code, '.htm', sep = '')
  tbls <- read_html(url) %>% 
    gsub(pattern='<!--', replacement='') %>% 
    gsub(pattern='-->', replacement='') %>%
    readHTMLTable()
  
  # Extract tables of interest into local data frames and prepare for merging
  offense <- scrape_table(tbls, 'player_offense', game_code, names)
  passing <- scrape_table(tbls, 'passing_advanced', game_code, names) %>%
    select(-c('Team', 'Cmp', 'PassAtt', 'PassYds', 'Sk'))
  rushing <- scrape_table(tbls, 'rushing_advanced', game_code, names) %>%
    select(-c('Team', 'RushAtt', 'RushYds'))
  receiving <- scrape_table(tbls, 'receiving_advanced', game_code, names) %>%
    select(-c('Team', 'Tgt', 'Rec', 'RecYds', 'RecTD'))
  
  # Merge data frames together
  join <- c('Player', 'Game')
  temp <- merge(merge(merge(offense, passing, by = join, all.x = TRUE), rushing, by = join, all.x = TRUE), receiving, by = join, all.x = TRUE)
  names(temp) <- names[['master']]
  
  return(temp)
}
```

A helper function for scrape_tables() that individually scrapes, renames, tags, and filters the tables of interest. 
```{r}
scrape_table <- function(tbls, table, game_code, names) {
  temp <- data.frame(tbls[[table]])
  
  # Rename, tag, and filter data frame
  names(temp) <- names[[table]]
  temp <- temp %>%
    mutate(Game = game_code) %>%
    filter(Player != '', Player != 'Player')
  
  return(temp)
}
```

# Scraping

Scrape tables of interest into merged master data frame and export to local .csv file. 
```{r}
master <- scrape_stats(years, names)

master %>%
  write.csv(paste('./Data/masterDF[', years[1], '-', years[length(years)], '].csv', sep = ''))
```



